%!TEX root = dissertation.tex
\addpublication{Zheng Zhao, Toni Karvonen, Roland Hostettler, and Simo S\"{a}rkk\"{a}}{Taylor moment expansion for continuous-discrete Gaussian filtering}{IEEE Transactions on Automatic Control}{Volume 66, Issue 9, Pages 4460--4467}{December}{2020}{Zheng Zhao, Toni Karvonen, Roland Hostettler, and Simo S\"{a}rkk\"{a}}{paperTME}
\addcontribution{Zheng Zhao wrote the article and produced the results. The stability analysis is mainly due to Toni Karvonen. Roland Hostettler gave useful comments. Simo S\"{a}rkk\"{a} contributed the idea.}
\adderrata{In Example 7, the coefficient $\Phi_{x,2}$ should multiply with a factor $2$.}

\addpublication{Zheng Zhao, Muhammad Emzir, and Simo S\"{a}rkk\"{a}}{Deep state-space Gaussian processes}{Statistics and Computing}{Volume 31, Issue 6, Article number 75, Pages 1--26}{September}{2021}{Zheng Zhao, Muhammad Emzir, and Simo S\"{a}rkk\"{a}}{paperSSDGP}
\addcontribution{Zheng Zhao wrote the article and produced the results. Muhammad Emzir and Simo S\"{a}rkk\"{a} gave useful comments.}

\addpublication{Zheng Zhao, Simo S\"{a}rkk\"{a}, and Ali Bahrami Rad}{Kalman-based spectro-temporal ECG analysis using deep convolutional networks for atrial fibrillation detection}{Journal of Signal Processing Systems}{Volume 92, Issue 7, Pages 621--636}{April}{2020}{Zheng Zhao, Simo S\"{a}rkk\"{a}, and Ali Bahrami Rad}{paperKFSECG}
\addcontribution{Zheng Zhao wrote the article and produced the results. Ali Bahrami Rad helped with the experiments. Simo S\"{a}rkk\"{a} came up with the spectro-temporal idea.}

\addpublication[conference]{Zheng Zhao, Filip Tronarp, Roland Hostettler, and Simo S\"{a}rkk\"{a}}{State-space Gaussian process for drift estimation in stochastic differential equations}{Proceedings of the 45th IEEE International Conference on Acoustics, Speech and Signal Processing (ICASSP)}{Barcelona, Spain, Pages 5295--5299}{May}{2020}{IEEE}{paperDRIFT}
\addcontribution{Zheng Zhao wrote the article and produced the results. Filip Tronarp provided codes for the iterated posterior linearisation filter. Roland Hostettler gave useful comments. Idea was due to Simo S\"{a}rkk\"{a}.}

\addpublication[conference]{Zheng Zhao, Simo S\"{a}rkk\"{a}, and Ali Bahrami Rad}{Spectro-temporal ECG analysis for atrial fibrillation detection}{Proceedings of the IEEE 28th International Workshop on Machine Learning for Signal Processing (MLSP)}{Aalborg, Denmark, 6 pages}{September}{2018}{IEEE}{paperKFSECGCONF}
\addcontribution{Zheng Zhao wrote the article and produced the results. Ali Bahrami Rad helped with the experiments. Simo S\"{a}rkk\"{a} came up with the spectro-temporal idea.}

\addpublication[accepted]{Sarang Thombre, Zheng Zhao, Henrik Ramm-Schmidt, Jos\'{e} M. Vallet García, Tuomo Malkam\"{a}ki, Sergey Nikolskiy, Toni Hammarberg, Hiski Nuortie, M. Zahidul H. Bhuiyan, Simo S\"{a}rkk\"{a}, and Ville V. Lehtola}{Sensors and AI techniques for situational awareness in autonomous ships: a review}{IEEE Transactions on Intelligent Transportation Systems}{20 pages}{September}{2020}{IEEE}{paperMARITIME}
\addcontribution{Zheng Zhao wrote the reviews of AI techniques and produced corresponding results.}

\addpublication[submitted]{Zheng Zhao, Rui Gao, and Simo S\"{a}rkk\"{a}}{Hierarchical Non-stationary temporal Gaussian processes with $L^1$-regularization}{Statistics and Computing}{20 pages}{May}{2021}{Zheng Zhao, Rui Gao, and Simo S\"{a}rkk\"{a}}{paperRNSSGP}
\addcontribution{Zheng Zhao wrote the article and produced the results. Rui Gao contributed the convergence analysis. Simo S\"{a}rkk\"{a} gave useful comments.}
